Detecting Treatment-Subgroup Interactions in Clustered Data with Generalized Linear Mixed-Effects Model Trees

### 2017-04-24 pkg/glmertree/paper/treatment_interactions.Rnw
Paper update
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@1272 255f4249-2263-4417-b1ea-8498779e27a7

<<simulation-data>>=
load("treespecs_study1.dat")
treespecs.long$N <- factor(as.numeric(as.character(treespecs.long$N)), ordered = TRUE)
treespecs.long$sigmab <- factor(as.numeric(as.character(treespecs.long$sigmab)), ordered = TRUE)
treespecs.long$treatdiff <- factor(as.numeric(as.character(treespecs.long$treatdiff)), ordered = TRUE)
treespecs.long$truetree.values <- as.numeric(treespecs.long$truetree.values)-1
levels(treespecs.long$corUb)[levels(treespecs.long$corUb)=="bi and splitting U correlated"] <- "b correlated with splitting U"
levels(treespecs.long$corUb)[levels(treespecs.long$corUb)=="bi and non-splitting U correlated"] <- "b correlated with non-splitting U"
levels(treespecs.long$corUb)[levels(treespecs.long$corUb)=="uncorrelated"] <- "b and U uncorrelated"
treespecs.long2 <- treespecs.long[!is.na(treespecs.long$treesize.values),]
treespecs.long2$treesize.ind <- factor(treespecs.long2$treesize.ind)
treespecs.long2$truetree.ind <- factor(treespecs.long2$truetree.ind)
treespecs.long2$truetree.values <- as.numeric(treespecs.long2$truetree.values)-1
@

@@

\begin{figure}[tb!]
\centering
<<treesize-nointeract-xyplot, fig=TRUE, width=9, height=7>>=
load("noint_treesizes_long.dat")
treesizes.long$sigmabm <- factor(as.numeric(as.character(treesizes.long$sigmabm)), ordered = TRUE)
treesizes.long$N <- factor(as.numeric(as.character(treesizes.long$N)), ordered = TRUE)
treesizes.long$treesize.valuesD <- as.numeric(treesizes.long$treesize.values>1)
levels(treesizes.long$treesize.ind)[levels(treesizes.long$treesize.ind)=="GLM tree"] <- "LM tree"
levels(treesizes.long$treesize.ind)[levels(treesizes.long$treesize.ind)=="GLMM tree"] <- "LMM tree"
#N, corUb & sigmab have strongest main and/or interaction effects
aggdata.size <- aggregate(formula=treesize.valuesD ~ treesize.ind + N + sigmabm + corUbm, 
FUN=mean, data=treesizes.long)
levels(aggdata.size$corUbm)[levels(aggdata.size$corUbm)=="b and random U correlated"] <- 
"b and U correlated"
levels(aggdata.size$corUbm)[levels(aggdata.size$corUbm)=="uncorrelated"] <- 
"b and U uncorrelated"
xyplot(treesize.valuesD ~ sigmabm | N + corUbm, data = aggdata.size, groups=treesize.ind, 
type="b", ylab="Type-I error", xlab=expression(sigma[b]), par.settings=standard.theme("pdf",color=F), 
auto.key=list(space="top", columns=2, title=" ", cex.title=1,lines=T, points=T))
@
\caption{\small Type-I error rate of LM and LMM trees. Rows represent dependence between random effects ($b$) and one of the partitioning variables $U_k$; columns represent sample size.}
\label{fig:xyplot_treesize_nointeract}
\end{figure}

@@

\begin{figure}[tb!]
\centering
<<correlation-pw-xyplot, fig=TRUE, width=9, height=7>>=
load("pw_treespecs_long.dat")
treespecs.long$N <- factor(as.numeric(as.character(treespecs.long$N)), ordered=T)
treespecs.long$sigmab <- factor(as.numeric(as.character(treespecs.long$sigmab)), ordered=T)
treespecs.long$type <- factor(treespecs.long$type, ordered = TRUE, levels = 
c("linear", "both", "piecewise"))
levels(treespecs.long$type) <- gsub("linear","continuous", levels(treespecs.long$type))
levels(treespecs.long$correlation.ind)[levels(treespecs.long$correlation.ind)=="GLMM"] <- "LMM"
levels(treespecs.long$correlation.ind)[levels(treespecs.long$correlation.ind)=="GLMM tree"] <- "LMM tree"
aggdata.cor <- aggregate(formula=correlation.values ~ correlation.ind + np + type + N, FUN=mean, 
data=treespecs.long)
xyplot(correlation.values ~ type | N + np, data = aggdata.cor, groups=correlation.ind, type="b",
ylab="correlation", xlab="interaction type", par.settings=standard.theme("pdf",color=F), 
auto.key=list(space="top", columns=2, title=" ", cex.title=1,lines=T, points=T),
scales=list(y=list(limits = c(-.1,1.1))))
@
\caption{\small Average predictive accuracy of LMMs and LMM trees. Predictive accuracy of trees is defined as the correlation between the true and predicted differences between Treatment~1 and~2. Columns represent sample size, rows represent the number of covariates.}
\label{fig:xyplot_correlation_pw}
\end{figure}

>>>>

<<simulation-data>>=
load("treespecs_studyI.dat")
treespecs.long$N <- factor(as.numeric(as.character(treespecs.long$N)), ordered = TRUE)
treespecs.long$sigmab <- factor(as.numeric(as.character(treespecs.long$sigmab)), ordered = TRUE)
treespecs.long$treatdiff <- factor(as.numeric(as.character(treespecs.long$treatdiff)), ordered = TRUE)
treespecs.long$truetree.values <- as.numeric(treespecs.long$truetree.values)-1
levels(treespecs.long$corUb)[levels(treespecs.long$corUb)=="bi and splitting U correlated"] <- "b correlated with splitting U"
levels(treespecs.long$corUb)[levels(treespecs.long$corUb)=="bi and non-splitting U correlated"] <- "b correlated with non-splitting U"
levels(treespecs.long$corUb)[levels(treespecs.long$corUb)=="uncorrelated"] <- "b and U uncorrelated"
treespecs.long2 <- treespecs.long[!is.na(treespecs.long$treesize.values),]
treespecs.long2$treesize.ind <- factor(treespecs.long2$treesize.ind)
treespecs.long2$truetree.ind <- factor(treespecs.long2$truetree.ind)
treespecs.long2$truetree.values <- as.numeric(treespecs.long2$truetree.values)-1
@

@@

\begin{figure}[tb!]
\centering
<<treesize-nointeract-xyplot, fig=TRUE, width=9, height=7>>=
load("treespecs_studyII.dat")
treesizes.long$sigmabm <- factor(as.numeric(as.character(treesizes.long$sigmabm)), ordered = TRUE)
treesizes.long$N <- factor(as.numeric(as.character(treesizes.long$N)), ordered = TRUE)
treesizes.long$treesize.valuesD <- as.numeric(treesizes.long$treesize.values>1)
levels(treesizes.long$treesize.ind)[levels(treesizes.long$treesize.ind)=="GLM tree"] <- "LM tree"
levels(treesizes.long$treesize.ind)[levels(treesizes.long$treesize.ind)=="GLMM tree"] <- "LMM tree"
#N, corUb & sigmab have strongest main and/or interaction effects
aggdata.size <- aggregate(formula=treesize.valuesD ~ treesize.ind + N + sigmabm + corUbm, 
FUN=mean, data=treesizes.long)
levels(aggdata.size$corUbm)[levels(aggdata.size$corUbm)=="b and random U correlated"] <- 
"b and U correlated"
levels(aggdata.size$corUbm)[levels(aggdata.size$corUbm)=="uncorrelated"] <- 
"b and U uncorrelated"
xyplot(treesize.valuesD ~ sigmabm | N + corUbm, data = aggdata.size, groups=treesize.ind, 
type="b", ylab="Type-I error", xlab=expression(sigma[b]), par.settings=standard.theme("pdf",color=F), 
auto.key=list(space="top", columns=2, title=" ", cex.title=1,lines=T, points=T))
@
\caption{\small Type-I error rate of LM and LMM trees. Rows represent dependence between random effects ($b$) and one of the partitioning variables $U_k$; columns represent sample size.}
\label{fig:xyplot_treesize_nointeract}
\end{figure}

@@

\begin{figure}[tb!]
\centering
<<correlation-pw-xyplot, fig=TRUE, width=9, height=7>>=
load("treespecs_studyIII.dat")
treespecs.long$N <- factor(as.numeric(as.character(treespecs.long$N)), ordered=T)
treespecs.long$sigmab <- factor(as.numeric(as.character(treespecs.long$sigmab)), ordered=T)
treespecs.long$type <- factor(treespecs.long$type, ordered = TRUE, levels = 
c("linear", "both", "piecewise"))
levels(treespecs.long$type) <- gsub("linear","continuous", levels(treespecs.long$type))
levels(treespecs.long$correlation.ind)[levels(treespecs.long$correlation.ind)=="GLMM"] <- "LMM"
levels(treespecs.long$correlation.ind)[levels(treespecs.long$correlation.ind)=="GLMM tree"] <- "LMM tree"
aggdata.cor <- aggregate(formula=correlation.values ~ correlation.ind + np + type + N, FUN=mean, 
data=treespecs.long)
xyplot(correlation.values ~ type | N + np, data = aggdata.cor, groups=correlation.ind, type="b",
ylab="correlation", xlab="interaction type", par.settings=standard.theme("pdf",color=F), 
auto.key=list(space="top", columns=2, title=" ", cex.title=1,lines=T, points=T),
scales=list(y=list(limits = c(-.1,1.1))))
@
\caption{\small Average predictive accuracy of LMMs and LMM trees. Predictive accuracy of trees is defined as the correlation between the true and predicted differences between Treatment~1 and~2. Columns represent sample size, rows represent the number of covariates.}
\label{fig:xyplot_correlation_pw}
\end{figure}
